Twang only analysis summary

### 2025-06-23 analysis.Rmd
readme and diagram

data <- readr::read_csv(file.path(s_root, "Gina's Project/df_Final_240909.csv")) |> 
  mutate(
    outcome_va_vi = outcome_VA_logMAR >= 0.3,
    outcome_va_blind = outcome_VA_logMAR >= 1,
    outcome_va_vi_binary = if_else(outcome_va_vi, 1, 0),
    outcome_va_blind_binary = if_else(outcome_va_blind, 1, 0),
    ever_lapse_cat = factor(person_ever_lapse),
    ever_lapse_binary = if_else(ever_lapse_cat == TRUE, 1, 0),
    MRN = dplyr::min_rank(e_mrn_deidentified),
    gender_cat = factor(gender),
    race_ethnic_cat = factor(race_ethnicity_gp),
    insurance_cat = factor(insurance_gp),
    age_cat = cut(age, breaks = c(0, 20, 45, 65, Inf), laebsl = c("0-20", "21-45", "46-65", "65+")),
    person_dr = case_when(
      No_DR == TRUE ~ 0,
      NPDR == "Mild" ~ 1,
      NPDR == "Moderate" ~ 1,
      NPDR == "Severe" ~ 1,
      PDR == "Present" ~ 2,
    ),
    person_dr_severity = case_when(
      No_DR == TRUE ~ 0,
      NPDR == "Mild" ~ 1,
      NPDR == "Moderate" ~ 2,
      NPDR == "Severe" ~ 3,
      PDR == "Present" ~ 4,
    ),
    person_ever_treat = case_when(
      other_inject == 0 & anti_VEGF == 0 & focal_laser_flag == 0 & PRP_flag == 0 ~ FALSE,
      other_inject == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | PRP_flag == 1 ~ TRUE
    ),
    glaucoma_bef_hitplus_cat = case_when(
      Glaucoma_bef_hitplus != "Present" | is.na(Glaucoma_bef_hitplus) ~ FALSE,
      Glaucoma_bef_hitplus == "Present" ~ TRUE
    ),
    glaucoma_after_hitplus_cat = case_when(
      Glaucoma_after_hitplus != "Present" | is.na(Glaucoma_after_hitplus) ~ FALSE,
      Glaucoma_after_hitplus == "Present" ~ TRUE
    ),
    otherretina_bef_hitplus_cat = case_when(
      Otherretina_bef_hitplus != "Present" | is.na(Otherretina_bef_hitplus) ~ FALSE,
      Otherretina_bef_hitplus == "Present" ~ TRUE
    ),
    glaucoma_after_hitplus_cat = case_when(
      Otherretina_after_hitplus != "Present" | is.na(Otherretina_after_hitplus) ~ FALSE,
      Otherretina_after_hitplus == "Present" ~ TRUE
    ),
    other_inject = factor(other_inject),
    anti_VEGF = factor(anti_VEGF),
    focal_laser_flag = factor(focal_laser_flag),
    PRP_flag = factor(PRP_flag),
    glaucoma_bef_hitplus_cat = factor(glaucoma_bef_hitplus_cat),
    otherretina_bef_hitplus_cat = factor(otherretina_bef_hitplus_cat),
    catsurg_before_hitplus_cat = factor(cataract_surgery_bef_hitplus)
  )

@@

```{r get_m_data}
m_data <- data |> filter(
  !is.na(!CCI) & !is.na(DCSI)
)

@@

# Define a list of matching methods and parameters, each with a unique name
methods <- list(
  list(name = "nearest_glm", method = "nearest", distance = "glm", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "nearest_gam", method = "nearest", distance = "gam", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "nearest_gbm", method = "nearest", distance = "gbm", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "nearest_lasso", method = "nearest", distance = "lasso", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "nearest_rpart", method = "nearest", distance = "rpart", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "subclass_glm", method = "subclass", distance = "glm", subclasses = 5, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "nearest_mahalanobis", method = "nearest", distance = "mahalanobis", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "cbps", use="weightit", method="cbps", estimand="ATT", rerun_model=FALSE, rerun_diagnostics = FALSE),
  list(name = "entropy", use="weightit", method="ebal", estimand="ATT", rerun_model=FALSE, rerun_diagnostics = FALSE),
  list(name = "bart", use="weightit", method="bart", estimand="ATT", rerun_model=FALSE, rerun_diagnostics = FALSE),
  list(name = "twang_gbm", use="twanng", method = "twang", estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE)
)

>>>>

data <- readr::read_csv(file.path(s_root, "Gina's Project/df_Final_240909.csv")) |> 
  mutate(
    outcome_va_vi = outcome_VA_logMAR >= 0.3,
    outcome_va_blind = outcome_VA_logMAR >= 1,
    outcome_va_vi_binary = if_else(outcome_va_vi, 1, 0),
    outcome_va_blind_binary = if_else(outcome_va_blind, 1, 0),
    ever_lapse_cat = factor(person_ever_lapse),
    ever_lapse_binary = if_else(ever_lapse_cat == TRUE, 1, 0),
    MRN = dplyr::min_rank(e_mrn_deidentified),
    gender_cat = factor(gender),
    race_ethnic_cat = factor(race_ethnicity_gp),
    insurance_cat = factor(insurance_gp),
    age_cat = cut(age, breaks = c(0, 20, 45, 65, Inf), labels = c("0-20", "21-45", "46-65", "65+")),
    person_dr = case_when(
      No_DR == TRUE ~ 0,
      NPDR == "Mild" ~ 1,
      NPDR == "Moderate" ~ 1,
      NPDR == "Severe" ~ 1,
      PDR == "Present" ~ 2,
    ),
    person_dr_severity = case_when(
      No_DR == TRUE ~ 0,
      NPDR == "Mild" ~ 1,
      NPDR == "Moderate" ~ 2,
      NPDR == "Severe" ~ 3,
      PDR == "Present" ~ 4,
    ),
    person_ever_treat = case_when(
      other_inject == 0 & anti_VEGF == 0 & focal_laser_flag == 0 & PRP_flag == 0 ~ FALSE,
      other_inject == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | PRP_flag == 1 ~ TRUE
    ),
    glaucoma_bef_hitplus_cat = case_when(
      Glaucoma_bef_hitplus != "Present" | is.na(Glaucoma_bef_hitplus) ~ FALSE,
      Glaucoma_bef_hitplus == "Present" ~ TRUE
    ),
    glaucoma_after_hitplus_cat = case_when(
      Glaucoma_after_hitplus != "Present" | is.na(Glaucoma_after_hitplus) ~ FALSE,
      Glaucoma_after_hitplus == "Present" ~ TRUE
    ),
    otherretina_bef_hitplus_cat = case_when(
      Otherretina_bef_hitplus != "Present" | is.na(Otherretina_bef_hitplus) ~ FALSE,
      Otherretina_bef_hitplus == "Present" ~ TRUE
    ),
    otherretina_after_hitplus_cat = case_when(
      Otherretina_after_hitplus != "Present" | is.na(Otherretina_after_hitplus) ~ FALSE,
      Otherretina_after_hitplus == "Present" ~ TRUE
    ),
    other_inject = factor(other_inject),
    anti_VEGF = factor(anti_VEGF),
    focal_laser_flag = factor(focal_laser_flag),
    PRP_flag = factor(PRP_flag),
    glaucoma_bef_hitplus_cat = factor(glaucoma_bef_hitplus_cat),
    otherretina_bef_hitplus_cat = factor(otherretina_bef_hitplus_cat),
    catsurg_before_hitplus_cat = factor(cataract_surgery_bef_hitplus)
  )

@@

```{r get_m_data}
m_data <- data |> filter(
  !is.na(CCI) & !is.na(DCSI)
)

@@

# Define a list of matching methods and parameters, each with a unique name
methods <- list(
  list(name = "nearest_glm", method = "nearest", distance = "glm", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "nearest_gam", method = "nearest", distance = "gam", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "nearest_gbm", method = "nearest", distance = "gbm", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "nearest_lasso", method = "nearest", distance = "lasso", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "nearest_rpart", method = "nearest", distance = "rpart", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "subclass_glm", method = "subclass", distance = "glm", subclasses = 5, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "nearest_mahalanobis", method = "nearest", distance = "mahalanobis", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "cbps", use="weightit", method="cbps", estimand="ATT", rerun_model=FALSE, rerun_diagnostics = FALSE),
  list(name = "entropy", use="weightit", method="ebal", estimand="ATT", rerun_model=FALSE, rerun_diagnostics = FALSE),
  list(name = "bart", use="weightit", method="bart", estimand="ATT", rerun_model=FALSE, rerun_diagnostics = FALSE),
  list(name = "twang_gbm", use="twang", method = "twang", estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE)
)

### 2025-06-23 twang-analysis.Rmd
readme and diagram

>>>>

```{r forest_plots}
library(ggplot2)
library(dplyr)

# Extract marginal effects data from all calculated objects
# Each mfx object has columns: factor, AME, SE, z, p, lower, upper
forest_data <- bind_rows(
  # Overall population
  all_mfx |> 
    filter(factor == "ever_lapse_binary") |> 
    mutate(
      label = "Overall Population",
      group_type = "overall",
      n = nrow(m_data)
    ),

  # No DR or NPDR subgroups
  mfx_npdr_nodr_no_tx |> 
    mutate(
      label = "No DR or NPDR, no treatment",
      group_type = "subgroup",
      n = n_npdr_nodr_no_tx
    ),
  mfx_npdr_nodr_any_tx |> 
    mutate(
      label = "No DR or NPDR, any treatment",
      group_type = "subgroup", 
      n = n_npdr_nodr_any_tx
    ),

  # NPDR overall
  mfx_npdr_no_tx |> 
    mutate(
      label = "NPDR, no treatment",
      group_type = "subgroup",
      n = n_npdr_no_tx
    ),
  mfx_npdr_any_tx |> 
    mutate(
      label = "NPDR, any treatment",
      group_type = "subgroup",
      n = n_npdr_any_tx
    ),

  # NPDR by severity
  mfx_npdr_mild_no_tx |> 
    mutate(
      label = "Mild NPDR, no treatment",
      group_type = "subgroup",
      n = n_npdr_mild_no_tx
    ),
  mfx_npdr_mild_any_tx |> 
    mutate(
      label = "Mild NPDR, any treatment",
      group_type = "subgroup",
      n = n_npdr_mild_any_tx
    ),
  mfx_npdr_moderate_no_tx |> 
    mutate(
      label = "Moderate NPDR, no treatment",
      group_type = "subgroup",
      n = n_npdr_moderate_no_tx
    ),
  mfx_npdr_moderate_any_tx |> 
    mutate(
      label = "Moderate NPDR, any treatment",
      group_type = "subgroup",
      n = n_npdr_moderate_any_tx
    ),
  mfx_npdr_severe_no_tx |> 
    mutate(
      label = "Severe NPDR, no treatment",
      group_type = "subgroup",
      n = n_npdr_severe_no_tx
    ),
  mfx_npdr_severe_any_tx |> 
    mutate(
      label = "Severe NPDR, any treatment",
      group_type = "subgroup",
      n = n_npdr_severe_any_tx
    ),

  # PDR subgroups
  mfx_any_tx |> 
    mutate(
      label = "PDR, any treatment",
      group_type = "subgroup",
      n = n_any_tx
    ),
  mfx_anti_vegf |> 
    mutate(
      label = "PDR, on anti-VEGF only",
      group_type = "subgroup",
      n = n_anti_vegf
    ),
  mfx_pdr_prp |> 
    mutate(
      label = "PDR, on PRP only",
      group_type = "subgroup",
      n = n_pdr_prp
    )
) |>
  mutate(
    # Add sample sizes to labels
    label_with_n = paste0(label, " (n=", n, ")"),
    # Calculate confidence intervals
    Lower = AME - 1.96 * SE,
    Upper = AME + 1.96 * SE
  )

# Define order for plots (reversed to match original plot)
subgroup_order <- rev(c(
  "No DR or NPDR, no treatment", "No DR or NPDR, any treatment", "NPDR, no treatment",
  "NPDR, any treatment", "Mild NPDR, no treatment", "Mild NPDR, any treatment",
  "Moderate NPDR, no treatment", "Moderate NPDR, any treatment",
  "Severe NPDR, no treatment", "Severe NPDR, any treatment",
  "PDR, any treatment", "PDR, on anti-VEGF only", "PDR, on PRP only"
))

# Version 1: Subgroups only (matching original forest plot)
data_subgroups <- forest_data |> 
  filter(group_type == "subgroup") |>
  mutate(Labels = factor(label, levels = subgroup_order))

plot_subgroups <- ggplot(data_subgroups, aes(x = AME, y = Labels)) +
  geom_vline(xintercept = 0, color = "#be123c", linetype = "solid", size = 0.5) +
  geom_point(size = 3, color = "#2563eb") +
  geom_errorbarh(aes(xmin = Lower, xmax = Upper), height = 0.2, color = "#2563eb") +
  geom_text(
    aes(x = Lower, label = sprintf("%.3f", Lower)), 
    hjust = 1.2, size = 3, color = "black"
  ) +
  geom_text(
    aes(x = Upper, label = sprintf("%.3f", Upper)), 
    hjust = -0.2, size = 3, color = "black"
  ) +
  geom_text(
    aes(label = sprintf("%.3f", AME)), 
    hjust = -0.2, vjust = -1, size = 3, color = "black"
  ) +
  theme_minimal() +
  scale_x_continuous(
    limits = c(-0.000, 0.06),
    breaks = scales::pretty_breaks(n = 6)
  ) +
  labs(
    title = "The Average Marginal Effects of Lapse Across Selected Groups",
    x = "Average Marginal Effect (AME)",
    y = "Treatment Group",
    caption = paste0("All p-values < ", max(data_subgroups$p, na.rm = TRUE))
  ) +
  theme(
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10), 
    axis.title.y = element_text(size = 12, face = "bold", margin = margin(r = 20)),
    axis.title.x = element_text(size = 12, face = "bold", margin = margin(t = 20)),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5, margin = margin(b = 20)),
    plot.margin = margin(10, 20, 30, 20),
    panel.grid.major.y = element_line(color = "gray", linetype = "dotted"),
    plot.caption = element_text(size = 10, face = "italic")
  )

# Version 2: With overall population included
all_order <- c("Overall Population", subgroup_order)
data_all <- forest_data |>
  mutate(
    Labels = factor(label, levels = rev(all_order)),
    # Different colors for overall vs subgroups
    point_color = ifelse(group_type == "overall", "#dc2626", "#2563eb"),
    point_size = ifelse(group_type == "overall", 4, 3)
  )

plot_with_overall <- ggplot(data_all, aes(x = AME, y = Labels)) +
  geom_vline(xintercept = 0, color = "#be123c", linetype = "solid", size = 0.5) +
  # Add horizontal separator line after overall population
  geom_hline(yintercept = length(subgroup_order) + 0.5, color = "gray50", linetype = "dashed", alpha = 0.5) +
  geom_point(aes(color = point_color, size = point_size)) +
  geom_errorbarh(aes(xmin = Lower, xmax = Upper, color = point_color), height = 0.2) +
  geom_text(
    aes(x = Lower, label = sprintf("%.3f", Lower)), 
    hjust = 1.2, size = 3, color = "black"
  ) +
  geom_text(
    aes(x = Upper, label = sprintf("%.3f", Upper)), 
    hjust = -0.2, size = 3, color = "black"
  ) +
  geom_text(
    aes(label = sprintf("%.3f", AME)), 
    hjust = -0.2, vjust = -1, size = 3, color = "black"
  ) +
  scale_color_identity() +
  scale_size_identity() +
  theme_minimal() +
  scale_x_continuous(
    limits = c(-0.000, 0.06),
    breaks = scales::pretty_breaks(n = 6)
  ) +
  labs(
    title = "The Average Marginal Effects of Lapse Across All Groups",
    x = "Average Marginal Effect (AME)",
    y = "Treatment Group",
    caption = paste0("All p-values < ", max(data_all$p, na.rm = TRUE), 
                    "\nOverall population shown in red, subgroups in blue")
  ) +
  theme(
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10), 
    axis.title.y = element_text(size = 12, face = "bold", margin = margin(r = 20)),
    axis.title.x = element_text(size = 12, face = "bold", margin = margin(t = 20)),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5, margin = margin(b = 20)),
    plot.margin = margin(10, 20, 30, 20),
    panel.grid.major.y = element_line(color = "gray", linetype = "dotted"),
    plot.caption = element_text(size = 10, face = "italic")
  )

# Display the plots
print(plot_subgroups)
print(plot_with_overall)

# Save the plots
ggsave("forest_plot_subgroups_only.png", plot_subgroups, 
       width = 10, height = 8, dpi = 300, bg = "white")
ggsave("forest_plot_with_overall.png", plot_with_overall, 
       width = 10, height = 8.5, dpi = 300, bg = "white")

# Also save extracted data for reference
write.csv(forest_data, "forest_plot_data.csv", row.names = FALSE)
